---
title: "Data_607_Week3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# #1
Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

Load .csv file
```{r Load Data}

df = data.frame(read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"))
head(df)
```

Search for College Majors containing "DATA" or "STATISTICS"

```{r Find Matches}
lst <- df$Major[grepl("DATA",df$Major,fixed = TRUE) | grepl("STATISTICS",df$Major,fixed = TRUE) ]
lst
```

# #2
Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```{r Data Transformation}

library(tidyverse)
library(dplyr)

data <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"'

data <- str_extract_all(data, '([a-zA-Z ]+[a-zA-Z])')
data <- unlist(data)
data
```

# #3

Write in words, what the expressions will match.

 "(.)\\\\1\\\\1"
The expression will match any character that appears three times in a row. (i.e fff)


 "(.)(.)\\\\2\\\\1"
The expression will match the first character and the second character and then the second character followed by the first character. (i.e. 0110)


 "(..)\\\\1"
The expression will match the a pair of characters that repeat immediately (i.e oooo, ieie). 


 "(.).\\\\1.\\\\1"
The expression will match a character that appears three times will a different character between each repeat. (i.e. c%c$c)


 "(.)(.)(.).*\\\\3\\\\2\\\\1"
The expression will match the 1st, 2nd, and 3rd characters and then anything, followed by the 3rd, 2nd, and 1st characters. (i.e. abccba, abc0000cba)

# #4

Construct regular expressions to match words that:

Start and end with the same character.
Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

```{r}

library(tidyverse)

f <- c("sings", "church", "eleven")

str_view(f, "^(.).*\\1$", match = TRUE)


str_view(f, "(..).*\\1", match = TRUE)


str_view(f,"(.).*\\1.*\\1", match = TRUE)
```

