---
title: "Data_607_week5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Data: https://github.com/acatlin/data/blob/master/israeli_vaccination_data_analysis.xlsx, https://www.covid-datascience.com/post/israeli-data-how-can-efficacy-vs-severe-disease-be-strong-when-60-of-hospitalized-are-vaccinated
  

#Data Cleanup

The original excel file was loaded into R and made into a data frame. A wide data set format was created to transform the data. All data regarding an age group will be contained in the same row.The data for each column was extracted from the data frame via direct element addresses and the grepl command. The columns were added in the order of the data set format and an "Efficacy" column was created to contain the Efficacy v. Disease calculations. The final csv file can be found at: 


## August

```{r, message=FALSE}
library(tidyverse)
library(dplyr)
library(readxl)
df <- data.frame(read_excel("israeli_vaccination_data_analysis_start.xlsx"))
df
colnames(df)
```

```{r Age}
age <- df$Age[grepl("50", df$Age)]
```

```{r Population}
pop_nv <- c(as.integer(df$Population..[2]), as.integer(df$Population..[4]))
pop_fv <- c(as.integer(df$...3[2]), as.integer(df$...3[4]))

pop_nv_p <- c(as.double(df$Population..[3])*100, as.double(df$Population..[5])*100)
pop_fv_p <- c(as.double(df$...3[3])*100, as.double(df$...3[5])*100)
```

```{r Severe Cases}
nv_s <- c(as.integer(df$Severe.Cases[2]), as.integer(df$Severe.Cases[4]))
fv_s <- c(as.integer(df$...5[2]), as.integer(df$...5[4]))
```

```{r Israel Vaccination}
library(dplyr)
iv <- data.frame(age, pop_nv, pop_nv_p, pop_fv, pop_fv_p, nv_s, fv_s)

iv[nrow(iv)+1,] <- c("Total",sum(iv$pop_nv),18.2,sum(iv$pop_fv),78.7,
                                                     sum(iv$nv_s),sum(iv$fv_s))
iv[2:7] <- lapply(iv[2:7], function(x) as.numeric(as.character(x)))
iv
```

```{r}
iv <- iv %>% mutate(nv_s_p = nv_s*100000/pop_nv, fv_s_p = fv_s*100000/pop_fv, Efficacy = (1-(fv_s_p/nv_s_p))*100)
iv
```


```{r csv}
write.csv(iv, "Israel_Vaccination.csv")
```

#Israel Vaccination Background:

  Israel's total population is estimated to be at 9.053 million.
  
  Anyone 12 years and older is eligible for the Covid-19 vaccine.
  
  The requirement to be considered fully vaccinated has changed from 2 doses to 3 for the Pfizer vaccine, as of August 29, 2021. Israel's department of Health announced that it would offer the third booster shot of the vaccine to individuals 12 years or older. Studies from Israel have shown that individuals with the third dose are more than tenfold less likely to get severe disease. 
  
  Currently the Moderna and Pfizer vaccines are being administered in Israel.
  
https://www.cnbc.com/2021/08/30/israel-doubles-down-on-covid-booster-shots-as-breakthrough-cases-rise.html



#Analysis

#(1) Do you have enough information to calculate the total population? What does this total population represent?

```{r}
iv %>% select(age, pop_nv, pop_fv)
```

We have enough information to calculate the total population. The total has been calculated for both the unvaccinated and vaccinated population by simply adding the numbers from each group. The total for unvacinated individuals is 1,302,912 and the total for vaccinated individuals is 563,634.


The total population would represent all individuals that are eligible to receive the Covid-19 vaccine (12 yrs and older).


#(2) Calculate the Efficacy vs. Disease; Explain your results.	

 1 - (% fully vaxed severe cases per 100K / % not vaxed severe cases per 100K)
  
```{r}
iv %>% select(age, Efficacy)
```
 


#(3) From your calculation of efficacy vs. disease, are you able to compare the rate of severe cases in unvaccinated individuals to that in vaccinated individuals?	
```{r}
iv %>% select(age, nv_s, nv_s_p, fv_s, fv_s_p, Efficacy) 
rate_unvax <- iv$nv_s_p[2]/iv$nv_s_p[1]
rate_vax <- iv$fv_s_p[2]/iv$fv_s_p[1]

data.frame(rate_unvax, rate_vax)
```

The efficacy of the vaccines in both groups (50 or younger, older than 50) is high (80%+). The rate of severe cases among unvaccinated individuals is higher than the rate of severe cases among vaccinated individuals. The rates of severe cases are higher for both vaccinated and unvaccinated older individuals. The efficacy calculations show that the efficacy of the vaccine is slightly lower for older individuals. 


