---
title: "New_York_Times"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# NYT Archive API

Using the New York Times Archive API, we will be analyzing the articles from November of 2020, the election month for the 2020 election, in order to gauge its weight in the new industry.

```{r, message=FALSE}
library(jsonlite)
library(tidyverse)
library(wordcloud)
```

```{r, include=FALSE}
key = "O0n35AS6kl66lBmqBAvcTEa7GSoIhHqW"
```

```{r}
url = "https://api.nytimes.com/svc/archive/v1/2020/11.json?api-key="
data <- as.data.frame(fromJSON(paste(url,key,sep = "")))
```


```{r, }
d <- data %>% select(Abstract = response.docs.abstract, Snippet = response.docs.snippet, Lead_Paragraph = response.docs.lead_paragraph, response.docs.type_of_material, Section = response.docs.section_name, Subsection = response.docs.subsection_name, Doc_Type = response.docs.document_type)
d["Headline"] <- data$response.docs.headline$main
n <- nrow(d)
```

# What are the top 5 sections, and subsections
```{r}
data.frame(table(d$Section)) %>% mutate(Proportion = Freq/n)  %>% arrange(desc(Proportion)) %>% head(5)
data.frame(table(d$Subsection)) %>% mutate(Proportion = Freq/n)  %>% arrange(desc(Proportion)) %>% head(5)
```
The top five sections are: U.S., World, Opinion, Business Day, and Arts.
The top five sub sections are: Elections, Politics, Europe, Book Review, and Television.
Of all the New York Times articles 34.8% were related to the U.S.

# Election

As we are interested in the 2020 election, the data will be filtered to contain rows that have any of the top three (U.S., World, Opinion) as the section.

```{r}
election <- d %>% filter(Section == "U.S." | Section == "World" | Section == "Opinion")
head(election,3)
```
The "Headlines" function take in a word and returns its count in the "Headlines" column in the "elections" data frame. This function will be used to acquire the frequency of all the desired words in the word bank.

```{r}

Headlines <- function(x){

  counts <- data.frame(table(unlist(strsplit(tolower(election$Headline), " ")))) %>% arrange(desc(Freq))
  x <- counts[which(counts$Var1 == x),]
  return(x)

}
```

A word bank is created consisting of keywords related to the 2020 election. These words will be the targets for our analysis. We would like to see the frequency of their appearance in Article headlines.

```{r}

words <- c("biden", "trump",  "harris", "election", "2020", "vote", "president", "vice", "coronavirus", "covid-19", "american", "democrat", "republican", "swing", "state", "votes", "results", "fraud", "battleground", "progressive", "ballot" )

results <- c()
counter <- 1
for(word in words)
{
  results[[counter]] <- Headlines(word)
  counter <- counter +1
}
results <- as.data.frame(bind_rows(results))
results <- cbind(as.data.frame(words), bind_cols(words, results))
results <- results[-2]
results <- results[-2]
colnames(results) <- c("Word", "Count")

results %>% arrange(desc(Count))
```
From the analysis we can see that of the 2,644 articles dealing with the US, politics, and Opinion, the word "Election" appeared 745 times, followed by "Trump" (233) and "Biden" (205)

```{r}
wordcloud(words = results$Word, freq = results$Count,  max.words=200, random.order=FALSE, rot.per=0.35,            colors=brewer.pal(8, "Dark2"))
```

# Conclusion

With the results we have gotten, we can say that the 2020 election was an important event as it relates to news for November of 2020. The word with the highest could was "election", with the names of the two presidental candidates following behind. As a continuation of this work, we would like to create a more neutral word bank and apply it to the articles published in the election month of the past few presidental elections.


